---
title: "Bank EDA"
author: "Megan Riley"
date: "3/13/2020"
output: pdf_document
---

```{r Setup}
library(readr)
library(tidyverse)
library(MASS)
library(ggplot2)
library(dplyr)
library(here)

root = here()


bank_20 = read.csv(paste(root,"/data/bank-additional/bank-additional-full.csv", sep = ""), sep=";")

bank_17 = read.csv(paste(root,"/data/bank/bank-full.csv", sep = ""), sep = ";")



```

## Summary

Unknown whether we should work with both data sets or if Dr. Turner is good with us choosing one. My vote is for bank_20 if we can choose. 

Variable Notes: 
-Duration is a variable not known until Y is determined, duration is the duration of the call when attempting to sell the term deposit package.  
- No NAs, uses unknown in places otherwise
-Campaign is  # of contacts, minimum 1 b/c it includes this contact in the data, even if hte contact was unsuccessful
-pdays needs to be potentially cleaned where 999 should equal NA or potentially switched to a categorical variable
-Do not understand some of the later variables, seem to be more socially based. 


```{r}
#Dr Turner is heavily requesting summary stats
summary(bank_20)

#Does not look like any NAs in either data set
sapply(bank_20, function(x) sum(is.na(x)))
sapply(bank_17, function(x) sum(is.na(x)))

#Set any predictions without using Duration or Y


#pdays- about 40k of the 41k are at level 999, no previous contact 
#could bin this data 
hist(bank_20$pdays)

temp = bank_20 %>% filter(pdays != 999)
dim(temp)
hist(temp$pdays)

```

## Uneven Split in outcomes

Yes happens about 10% of the time, where no is the response 90% of the time. This unbalance makes it difficult to predict. 
- Can balance the train/test split, what else have we learned about predicting unbalanced outcomes? 


```{r}

```
