---
title: "Bank EDA"
author: "Megan Riley"
output: pdf_document
---

```{r Setup}
library(readr)
library(tidyverse)
library(MASS)
library(ggplot2)
library(dplyr)
library(here)
library(GGally)
library(ResourceSelection)
library(car)
library(caret)

root = here()


bank_20 = read.csv(paste(root,"/data/bank-additional/bank-additional-full.csv", sep = ""), sep=";")

bank_17 = read.csv(paste(root,"/data/bank/bank-full.csv", sep = ""), sep = ";")



```

## Summary

Unknown whether we should work with both data sets or if Dr. Turner is good with us choosing one. My vote is for bank_20 if we can choose. 

Variable Notes: 
-Duration is a variable not known until Y is determined, duration is the duration of the call when attempting to sell the term deposit package.  
- No NAs, uses unknown in places otherwise
-Campaign is  # of contacts, minimum 1 b/c it includes this contact in the data, even if hte contact was unsuccessful
-pdays needs to be potentially cleaned where 999 should equal NA or potentially switched to a categorical variable
-Do not understand some of the later variables, seem to be more socially based. 


```{r summary} 
#Dr Turner is heavily requesting summary stats
summary(bank_20)

#Does not look like any NAs in either data set
sapply(bank_20, function(x) sum(is.na(x)))
sapply(bank_17, function(x) sum(is.na(x)))

#Set any predictions without using Duration or Y


```


## Data Preprocessing

Pdays is a major variable that needs correcting, the large inclusion of the 999 category requires either splitting the variable or turning it into a categorical with bins. The first option will be used to create one boolean variable that indicates whether a customer has been contacted before(where the 999 code indicates they have not been) and another variable with 


```{r cleaning}

clean_bank_20 = as.data.frame(bank_20)
#pdays- about 40k of the 41k are at level 999, no previous contact 
#could bin this data 
hist(bank_20$pdays)

temp = bank_20 %>% filter(pdays != 999)
dim(temp)
hist(temp$pdays)
summary(temp$pdays)

#within 5 days, 10 , 15, 30 and never

clean_bank_20$newpdays = case_when(bank_20$pdays == 999 ~ "Never",
                     bank_20$pdays >= 15 ~ "Within 30 Days", 
                     bank_20$pdays >= 10 & bank_20$pdays < 15 ~ "Within 15 Days",
                     bank_20$pdays >= 5 & bank_20$pdays < 10 ~ "Within 10 Days",
                     bank_20$pdays < 5 ~ "Within 5 Days") 

#clean_bank_20 = dplyr::select(clean_bank_20, -pdays) 


#Dr Turner's other suggestion 
#Set up a categorical variable to turn the continuous variable on or off. 
#any use of this would have to be both in tandem
#ie y ~ altpdays_cat*altpdays_cont
alt_pdays_cat = ifelse(bank_20$pdays == 999, 0, 1)
#remains the same as original pdays, 

alt_clean_bank_20 = bank_20
alt_clean_bank_20$pdays_cat = alt_pdays_cat


```

## EDA Plots of Specific Variables by Y Response


```{r EDA}

#Age 
clean_bank_20 %>% ggplot(aes(y= age,fill = y)) + geom_boxplot() + ggtitle("Distribution of Age by Status of Response")

#Job
clean_bank_20 %>% ggplot(aes(x = job, fill = y)) + geom_histogram(stat = "count")
clean_bank_20 %>% ggplot(aes(x = job, fill = y)) + geom_bar(position = "fill")


#Marital
clean_bank_20 %>% ggplot(aes(x = marital, fill = y)) + geom_histogram(stat = "count")
clean_bank_20 %>% ggplot(aes(x = marital, fill = y)) + geom_bar(position = "fill")



#Education
clean_bank_20 %>% ggplot(aes(x = education, fill = y)) + geom_histogram(stat = "count")
clean_bank_20 %>% ggplot(aes(x = education, fill = y)) + geom_bar(position = "fill")

#Default
clean_bank_20 %>% ggplot(aes(x = default, fill = y)) + geom_histogram(stat = "count")
clean_bank_20 %>% ggplot(aes(x = default, fill = y)) + geom_bar(position = "fill")


#Housing
clean_bank_20 %>% ggplot(aes(x = housing, fill = y)) + geom_histogram(stat = "count")
clean_bank_20 %>% ggplot(aes(x = housing, fill = y)) + geom_bar(position = "fill")


#Loan

clean_bank_20 %>% ggplot(aes(x = loan, fill = y)) + geom_histogram(stat = "count")
clean_bank_20 %>% ggplot(aes(x = loan, fill = y)) + geom_bar(position = "fill")

#Contact
clean_bank_20 %>% ggplot(aes(x = contact, fill = y)) + geom_histogram(stat = "count")
clean_bank_20 %>% ggplot(aes(x = contact, fill = y)) + geom_bar(position = "fill")


#Month
clean_bank_20 %>% ggplot(aes(x = month, fill = y)) + geom_histogram(stat = "count")
clean_bank_20 %>% ggplot(aes(x = month, fill = y)) + geom_bar(position = "fill")


#Day_of_week
clean_bank_20 %>% ggplot(aes(x = day_of_week, fill = y)) + geom_histogram(stat = "count")
clean_bank_20 %>% ggplot(aes(x = day_of_week, fill = y)) + geom_bar(position = "fill")


#duration

clean_bank_20 %>% ggplot(aes(y = duration, fill = y)) + geom_boxplot() + ggtitle("Duration by Status of Response")

clean_bank_20 %>% ggplot(aes(y = log(duration), fill = y)) + geom_boxplot() + ggtitle("Logged Duration by Status of Response")

#Campaign



#newPdays

clean_bank_20 %>% ggplot(aes(x =newpdays, fill = y)) + geom_histogram(stat = "count")
clean_bank_20 %>% ggplot(aes(x = newpdays, fill = y)) + geom_bar(position = "fill")


#Previous



#poutcome
clean_bank_20 %>% ggplot(aes(x =poutcome, fill = y)) + geom_histogram(stat = "count")
clean_bank_20 %>% ggplot(aes(x = poutcome, fill = y)) + geom_bar(position = "fill")


#emp.var.rate
clean_bank_20 %>% ggplot(aes(y = emp.var.rate, fill = y)) + geom_boxplot() + ggtitle("emp.var.rate by Status of Response")
boxplot(clean_bank_20$emp.var.rate)


#cons.price.idx
clean_bank_20 %>% ggplot(aes(y = cons.price.idx, fill = y)) + geom_boxplot() + ggtitle("CPI by Status of Response")

#something here potentially, but is messy
clean_bank_20 %>% ggplot(aes(x = month, y = cons.price.idx, fill = y)) + geom_boxplot() + ggtitle("CPI by month  by Status of Response")


#CCI and CPI

clean_bank_20 %>% ggplot(aes(x = cons.conf.idx, y = cons.price.idx, color = y)) + geom_jitter() + ggtitle("CPI by month  by Status of Response")


#cons.conf.idx
clean_bank_20 %>% ggplot(aes(y = cons.conf.idx, fill = y)) + geom_boxplot() + ggtitle("CCI by Status of Response")

clean_bank_20 %>% ggplot(aes(x = month, y = cons.conf.idx, color = y)) + geom_point() + ggtitle("CCI by month  by Status of Response")


clean_bank_20 %>% ggplot(aes(x = cons.conf.idx, fill = y)) + geom_histogram(stat = "count") + ggtitle("CCI by Status of Response")




#euribor3m

#nr.employed


#working with the ftables

#probs needs to be a factor
#ftable(addmargins(clean_bank_20$y, clean_bank_20$month))

```


## Multicolinearity and Interactions

We need to check our multicolinearity between predictor variables as well as interactions, for example, month information of a cyclical nature may be reflected in some of the economic indicators or dependent on them. 


```{r interactions}

#Multicoliniearity in the continuous variables. 
cont_bank = clean_bank_20[,c(1,11,12,13,15,16,17,18,19)]
ggcorr(cont_bank, label = TRUE, hjust = 1  )

###
#Interactions on Categorical + Categorical  Variables

#Job and Marital interaction  - Potentially Useful
clean_bank_20 %>% ggplot(aes(x = job, fill = marital))  + geom_bar(position = "fill")

#Job and DEfault - Worth keeping, some interaction
clean_bank_20 %>% ggplot(aes(x = job, fill = default))  + geom_bar(position = "fill")

#Pdays sole categorical var and contact,   A little interaction, 
clean_bank_20 %>% ggplot(aes(x = newpdays, fill = contact))  + geom_bar(position = "fill")
 
#education and job - Complex visually due to categories, but some interaction present for sure
clean_bank_20 %>% ggplot(aes(x = education, fill = job))  + geom_bar(position = "fill")

#


###
#Continuous + Categorical
#Job and nr. employed - Trend is confusing, but potentially some interaction
#Type of job related to employment in country? 
clean_bank_20 %>% ggplot(aes(x = job, y = nr.employed, fill = y)) + geom_boxplot()

#Cons confidence and job - not quite significant
clean_bank_20 %>% ggplot(aes(x = job, y = cons.conf.idx, fill = y)) + geom_boxplot()

#Month with cons conf and duration, month seems useful, but interactions are confusing, seem odd. 
clean_bank_20 %>% ggplot(aes(x = month, y = cons.conf.idx, fill = y)) + geom_boxplot()
clean_bank_20 %>% ggplot(aes(x = month, y = log(duration), fill = y)) + geom_boxplot()

#Original Pdays and contact - VV Hard to truly see interaction here dunno how to turn on/off for plot only quickly 
#Lets limit pdays interactions then..
clean_bank_20 %>% ggplot(aes(x = contact, y = pdays, fill = y)) + geom_boxplot()



#Contact Type and Duration - I Don't think this is significant
clean_bank_20 %>% ggplot(aes(x = contact, y = log(duration), fill = y)) + geom_boxplot()



###
#Continuous + Continous

#Use the ggcorr plot from before...
#Emp. var rate, cons price index,  and euribor3m are multicolinear - include 1 or none




#job*default + contact*duration + pdays*contact + pdays*duratio

```


## Predicting Duration, in order to predict calls. 

Basically if we know there is a strong relationship between duration and the result we are predicting, we can use information that explains duration to therefore explain response. Since it is invalid to use duratuion in my opinion.

- Is this valid? Possibly? 
- Is it helpful? Maybe? 

```{r }

duration_model = lm(duration ~ ., data = bank_20)

```

## Uneven Split in outcomes

Yes happens about 10% of the time, where no is the response 90% of the time. This unbalance makes it difficult to predict. 
- Can balance the train/test split, what else have we learned about predicting unbalanced outcomes? 



```{r splitData}
#Currently produces a train set of 52 n / 48 y
#90/10 yes train test split
set.seed(4567)

yes_indices = which(clean_bank_20$y == "yes")
yes_train_indices = sample(yes_indices, length(yes_indices) * .9)
no_indices = which(clean_bank_20$y == "no")
#
no_train_indices = sample(no_indices, length(yes_indices))
train_indices = c(no_train_indices,yes_train_indices)

balanced_train_bank_20 = clean_bank_20[train_indices,]

test_bank_20 = clean_bank_20[-train_indices,]
summary(balanced_train_bank_20$default)

```

## Simple Baseline Predictions

```{r}
#using a handful of vars I identified as probably useful on a first pass
simple.model = glm(y~ job + marital + month + newpdays + poutcome, data = balanced_train, family = binomial(link = "logit"))

#VIF of simple model

(vif(simple.model)[,3])^2

#Prediction of simple model
pred = ifelse(simple.model$fitted.values > .5, 1, 0)
predtable = table(pred, clean_bank_20$y )
predtable

#prediction of full model

full.model = glm(y~ age+ job + marital + education + default + housing + loan + contact
                 + month + day_of_week + campaign + previous + poutcome + emp.var.rate + 
                   cons.price.idx + cons.conf.idx + euribor3m + nr.employed + newpdays , data = clean_bank_20 , family = binomial(link="logit"))

fullpred = ifelse(full.model$fitted.values > .5, 1,0)
fullpredtable = table(fullpred, clean_bank_20$y)
fullpredtable

```

